Numerical simulation of the Cordilleran ice sheet \\ through the last glacial cycle

### 2014-06-26 kappa.tex
Release full outline with planned figs and tables.
Correct the ice rheology formula.

\maketitle
\linenumbers

% ======================================================================
\section{Introduction}
% ======================================================================
% Ice as viscous matter
% Glaciers and ice sheets
% The last glacial cycle

% Fig. crevasses
% Fig. glacier flowing
% Fig. recent climate evolution
% Fig. present and palaeo-ice sheets

% ======================================================================
\section{Field area description}
% ======================================================================
% Geographic setting
% Climatic setting
% Palaeo-glaciology

% Fig. map of North America
% Fig. map of the Cordillera?
% Fig. photos of the Cordillera
% Fig. photos of fieldwork
% Fig. Johan lineation map?

@@

% Fig. PISM = ice + bedrock

@@

@@

@@

Thirdly, after defining the strain-rate tensor
${\SRT = \frac{1}{2}(\tgrad{\vv} + \tgrad{\vv}^{\mathrm{T}})}$,
and the deviatoric stress tensor, ${\DST = \CST - p\,\IDT}$,
where $p=\frac{1}{3}\tr(\CST)$ is the hydrostatic pressure and
$\tens{\delta}$ is the identity tensor, a \emph{constitutive law} for ice based
on laboratory experiments is used to relate these two quantities. It reads:
\begin{equation}
    \label{eqn:glenslaw}
    \SRT = A\,\tau_e^{n-1}\,\DST \,.
\end{equation}
The equivalent stress, $\tau_e$, is defined by
${\tau_e}^2 = -\mathrm{II}_{\DST} = \frac{1}{2} \tr(\DST^2)$,
where $\mathrm{II}_{\DST}$ is the second invariant of the stress tensor.
The ice softness coefficient, $A$, depends on the deviation of temperature $T$
from the pressure-melting point $T_m$ through
\begin{equation}
    A = A_0\,e^\frac{-Q}{R(T-T_m)} \,.
\end{equation}

@@

% Fig. SIA + SSA

@@

The till cohesion $c_0$ is assumed to be zero. The effective pressure on the
till, $N$, is determined by the modelled amount of water at the bed. However,
two different parametrisations are used within this thesis. In {\CCLI}, using
PISM~0.5, the effective pressure is linearly related to water content by a
simple parametrisation,
\begin{equation}
    N = \rho gH (1 - \alpha \frac{W}{W_{max}}) \,.
\end{equation}

In {\CCYC}, using PISM~0.6, the effective pressure is determined by an
empirical parametrisation based on laboratory experiments on Antarctic till,
\begin{equation}
    N = \delta \rho gH \, 10^{(e_0/C_c) (1 - (W/W_{max}))} \,.
\end{equation}

@@

@@

@@

@@

% ======================================================================
\section{Software tools and contributions}
% ======================================================================
% Computational workflow
% Contributions to PISM
% PyPDD
% pismplotlib
% r.in.worldclim
% r.out.pism

% Fig. flow chart
% Fig. pypdd example
% Fig. pismplotlib example?

% ======================================================================
\section{Results summary}
% ======================================================================
% Constant-climate runs
% Effect of daily temperature variability
% Transient-climate runs

% Fig. 3km LGM quiver
% Fig. sigma distribution map
% Fig. sigma effect on Cordillera runs
% Fig. deglaciation streamplot
% Fig. cumulative ice displacement map

>>>>

\maketitle
\linenumbers
\tableofcontents

% ======================================================================
\section{Introduction}
% ======================================================================

\todo{Figure. Glacier crevasse photo}\\
\todo{Figure. Glacier flowing photo}\\
\todo{Figure. Glacier moraine photo}\\
\todo{Figure. Benthic oxygen isotopes stack}\\
\todo{Figure. Modern and former ice sheets}

% ======================================================================
\section{Field area description}
% ======================================================================

% ----------------------------------------------------------------------
\subsection{Geographic setting}
% ----------------------------------------------------------------------

\todo{Figure. Location map of North America}\\
\todo{Figure. Topographic map of the Cordillera}\\
\todo{Table. LCC projection parameters?}

% ----------------------------------------------------------------------
\subsection{Climatic setting}
% ----------------------------------------------------------------------

\todo{Figure. Cordillera wet climate photo}\\
\todo{Figure. Cordillera dry climate photo}\\
\todo{Figure. Climatic patters}

% ----------------------------------------------------------------------
\subsection{Palaeo-glaciology}
% ----------------------------------------------------------------------

\todo{Figure. Glacierized region photo}\\
\todo{Figure. Tuya Lake moraine photo}\\
\todo{Figure. Meltwater channels photo}\\
\todo{Figure. Lineation map}

@@

\todo{Figure. PISM interfaces}

@@

\todo{Table. model variables and parameters}

@@

\todo{Figure. Ice flow problem sketch}

@@

Thirdly, after defining the strain-rate tensor
${\SRT = \frac{1}{2}(\tgrad{\vv} + \tgrad{\vv}^{\mathrm{T}})}$,
and the deviatoric stress tensor, ${\DST = \CST - p\,\IDT}$,
where $p=\frac{1}{3}\tr(\CST)$ is the hydrostatic pressure and
$\tens{\delta}$ is the identity tensor, a \emph{constitutive law} for ice based
on laboratory experiments is used to relate these two quantities. It reads:
\begin{equation}
    \label{eqn:glenslaw}
    \SRT = A\,\tau_e^{n-1}\,\DST \,.
\end{equation}
The equivalent stress, $\tau_e$, is defined by
${\tau_e}^2 = -\mathrm{II}_{\DST} = \frac{1}{2} \tr(\DST^2)$,
where $\mathrm{II}_{\DST}$ is the second invariant of the stress tensor.
The ice softness coefficient, $A$, depends on ice temperature, pressure, and
water, $\omega$ through an Arrhenius-type law,
\begin{equation}
    A =
    \begin{cases}
        A_c (1+f\omega)\,e^\frac{-Q_c}{RT_{pa}}
            & \text{if}\ T < T_{crit} \,, \\
        A_w (1+f\omega)\,e^\frac{-Q_w}{RT_{pa}}
            & \text{if}\ T \ge T_{crit} \,,
    \end{cases}
\end{equation}
where $T_{pa}$ is the pressure-adjusted temperature calculated using the
Clapeyron relation, ${T_{pa} = T - \beta p}$.

@@

\todo{Figure. SIA and SSA}

@@

The till cohesion $c_0$ is assumed to be zero. The effective pressure on the
till, $N$, is determined by the modelled amount of water at the bed. However,
two different parametrisations are used within this thesis. In {\CCLI}, using
PISM~0.5, the effective pressure is linearly related to water content by a
simple parametrisation,
\begin{equation}
    N = \rho gh (1 - \alpha \frac{W}{W_{max}}) \,.
\end{equation}

In {\CCYC}, using PISM~0.6, the effective pressure is determined by an
empirical parametrisation based on laboratory experiments on Antarctic till,
\begin{equation}
    N = \delta \rho gh \, 10^{(e_0/C_c) (1 - (W/W_{max}))} \,.
\end{equation}

@@

\todo{Figure. SMB model algorithm}

@@

\todo{Figure. Boundary models data flow chart}

@@

\todo{Table. Data sources?}

@@

% ======================================================================
\section{Software tools and contributions}
% ======================================================================

% ----------------------------------------------------------------------
\subsection{Computational workflow}
% ----------------------------------------------------------------------
\todo{Figure. computational workflow}

% ----------------------------------------------------------------------
\subsection{A Python PDD model}
% ----------------------------------------------------------------------
\todo{Figure. PyPDD demo example}

% ----------------------------------------------------------------------
\subsection{Pre-processing scripts for GRASS GIS}
% ----------------------------------------------------------------------

% ----------------------------------------------------------------------
\subsection{Plotting tools for PISM}
% ----------------------------------------------------------------------
\todo{Figure. Plotting tools example}

% ----------------------------------------------------------------------
\subsection{Contributions to PISM}
% ----------------------------------------------------------------------
\todo{Table. PISM ``issues'' identified}

% ======================================================================
\section{Results summary}
% ======================================================================

% ----------------------------------------------------------------------
\subsection{Constant-climate runs}
% ----------------------------------------------------------------------
\todo{Figure. 3km last glacial maximum poster run?}

% ----------------------------------------------------------------------
\subsection{Daily temperature variability}
% ----------------------------------------------------------------------
\todo{Figure. Sigma distribution map}\\
\todo{Figure. Sigma effect on Cordillera runs}

% ----------------------------------------------------------------------
\subsection{Transient-climate runs}
% ----------------------------------------------------------------------
\todo{Figure. Deglaciation streamplot}\\
\todo{Figure. Cumulative ice displacement map}
